```{r setup}
#| message: false
#| include: false
library(tidyverse)
library(vroom)
library(hms)
```

# Cleaning {.unnumbered}

This chapter details all cleaning.

> Note: myClim does not do much cleaning <https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.14192>
>
> Note: SoilTemp database does not address data cleaning <https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.15123>

## Daily

```{r}
#| message: false
#| warning: false
raw <- vroom("outputs/tomst_data.tsv") %>% 
  gather(variable, value, -site, -date, -tomst_sensornum) %>% 
  group_by(site, variable, tomst_sensornum, date = as_date(date)) %>% 
  summarise(value = median(value))
cleaned <- raw %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  mutate(swc = ifelse(swc < 0, NA, swc)) %>% 
  mutate(swc = ifelse(tomst_sensornum == 94244926, NA, swc))
cleaned <- cleaned %>% 
  select(-t_air, -t_soil, -t_surface) %>% 
  left_join(cleaned %>% 
              group_by(site, tomst_sensornum) %>% 
              arrange(date) %>% 
              filter_at(c("t_soil", "t_air", "t_surface"), ~ abs(. - lag(.)) < 1.5) %>% 
              select(-swc)) %>% 
  gather(variable, value, -site, -date, -tomst_sensornum)
```

```{r}
#| eval: false
data <- vroom("outputs/tomst_data.tsv") %>% 
  gather(variable, value, -site, -date, -tomst_sensornum) %>% 
  mutate(day = as_date(date))
daily_cleaned <- cleaned %>% 
  na.omit() %>% 
  rename(day = date) %>% 
  select(site, tomst_sensornum, variable, day) %>% 
  left_join(data)
write_tsv(daily_cleaned, "outputs/tomst_daily_cleaned.tsv")
```

### Temperature

-   Doubtful start: P26:P30 (all Paracou), can be removed by suppressing day with a temperature delta above 1.5°C
-   Doubtful end: all end, remove last day, can be removed by suppressing day with a temperature delta above 1.5°C
-   Doubtful t_soil: N02, could be linked to calibration offset but not according to calibration results, could also be natural
-   Globally seems pretty clean

```{r}
#| message: false
#| warning: false
raw %>% 
  filter(variable != "swc") %>% 
  mutate(tomst_sensornum = gsub("942449", "", tomst_sensornum)) %>% 
  mutate(site = str_sub(site, 1, 1)) %>% 
  ggplot(aes(date, value, col = variable)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~ paste0(site, tomst_sensornum)) +
  theme(axis.title = element_blank(), legend.position = "bottom")  +
  ggtitle("Raw")
```

```{r}
#| message: false
#| warning: false
cleaned %>% 
  filter(variable != "swc") %>% 
  mutate(tomst_sensornum = gsub("942449", "", tomst_sensornum)) %>% 
  mutate(site = str_sub(site, 1, 1)) %>% 
  ggplot(aes(date, value, col = variable)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~ paste0(site, tomst_sensornum)) +
  theme(axis.title = element_blank(), legend.position = "bottom")  +
  ggtitle("Daily cleaned")
```

### SWC

-   Doubtful start:
    -   P28:P30:
        -   first day
        -   can be removed easily with swc\>0
-   All doubtful:
    -   P26:
        -   note: 23-10-16 *does note seem to measure humidity*
        -   To be totally removed
-   Doubtful period:
    -   N12
        -   note: *sensor unearthed* between 23-08-28 and 23-08-29
        -   data: from 2023-04-21 to 2023-08-29 to be removed
        -   can be removed easily with swc\>0

```{r}
#| message: false
#| warning: false
raw %>% 
  filter(variable == "swc") %>% 
  mutate(tomst_sensornum = gsub("942449", "", tomst_sensornum)) %>% 
   mutate(site = str_sub(site, 1, 1)) %>% 
  ggplot(aes(date, value)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~ paste0(site, tomst_sensornum)) +
  theme(axis.title = element_blank()) +
  ggtitle("Raw")
```

```{r}
#| message: false
#| warning: false
cleaned %>% 
  filter(variable == "swc") %>% 
  mutate(tomst_sensornum = gsub("942449", "", tomst_sensornum)) %>% 
  mutate(site = str_sub(site, 1, 1)) %>% 
  ggplot(aes(date, value)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~ paste0(site, tomst_sensornum)) +
  theme(axis.title = element_blank()) +
  ggtitle("Daily cleaned")
```

### Hourly

Apart from null swc that have be easily removed, everything looked okay with individual inspection. Some sensors shows air and surface but not soil temperature high inrcease during a specific time of the day every day (e.g. N03 around 11:00) that can be explained by direct light exposure but are not errors as tomst sensors are protected against direct radiations.

```{r}
#| eval: false
data <- vroom("outputs/tomst_daily_cleaned.tsv")
hourly_fig <- function(num, data, dest){
  t <- data %>% 
  filter(tomst_sensornum == num)
  g <- t %>% 
  mutate(day = as_date(date), hour = as_hms(date)) %>% 
  ggplot(aes(hour, value, col = day)) +
  geom_line(aes(group = as.character(day))) +
  theme_bw() +
  facet_wrap(~ variable, scales = "free_y") +
  scale_colour_viridis_c(guide = "none") +
  ggtitle(num, unique(t$site)) +
  theme(axis.title = element_blank())
  ggsave(filename = file.path(dest, paste0(num, ".png")),
         plot = g)
} 
for(i in unique(data$tomst_sensornum))
  hourly_fig(i, data, "outputs/hourly_cleaning")
```

```{r}
#| eval: false
library(data.table)
data <- fread("outputs/tomst_daily_cleaned.tsv")
data <- data[value > 0, ]
fwrite(data, "outputs/tomst_hourly_cleaned.tsv", sep = "\t")
```
