```{r setup}
#| message: false
#| include: false
library(tidyverse)
library(vroom)
library(hms)
```

# Exploration {.unnumbered}

This chapter details all data exploration.

## Day, dial and diurnal

```{r prep}
#| eval: false
dir.create("tmp")
# daily
vroom("outputs/tomst_hourly_cleaned.tsv") %>% 
  group_by(site, tomst_sensornum, variable, day) %>% 
  summarise(l = min(value), m = median(value), h = max(value)) %>% 
  mutate(type = "daily") %>% 
  rename(date = day) %>% 
  gather(metric, value, -site, -date, -tomst_sensornum, -variable, -type) %>% 
  vroom_write(file = "tmp/daily.tsv")
# dial
vroom("outputs/tomst_hourly_cleaned.tsv") %>% 
  filter(as_hms(date) >= "07:00:00", as_hms(date) < "19:00:00") %>% 
  group_by(site, tomst_sensornum, variable, day) %>% 
  summarise(l = min(value), m = median(value), h = max(value)) %>% 
  mutate(type = "dial") %>% 
  rename(date = day) %>% 
  gather(metric, value, -site, -date, -tomst_sensornum, -variable, -type) %>% 
  vroom_write("tmp/dial.tsv")
# diurnal
vroom("outputs/tomst_hourly_cleaned.tsv") %>% 
  filter(as_hms(date) < "07:00:00" | as_hms(date) >= "19:00:00") %>% 
  group_by(site, tomst_sensornum, variable, day) %>% 
  summarise(l = min(value), m = median(value), h = max(value)) %>% 
  mutate(type = "diurnal") %>% 
  rename(date = day) %>% 
  gather(metric, value, -site, -date, -tomst_sensornum, -variable, -type) %>% 
  vroom_write("tmp/diurnal.tsv")
# merge
bind_rows(
  vroom("tmp/daily.tsv"),
  vroom("tmp/dial.tsv"),
  vroom("tmp/diurnal.tsv")
) %>% 
  vroom_write("outputs/tomst_daily_cleaned.tsv")
unlink("tmp", recursive = TRUE)
```

```{r}
#| message: false
#| warning: false
read_tsv("outputs/tomst_daily_cleaned.tsv") %>% 
  pivot_wider(names_from = metric, values_from = value) %>% 
  ggplot(aes(date, m, col = paste(site, tomst_sensornum), fill = paste(site, tomst_sensornum))) +
  geom_ribbon(aes(ymin = l, ymax = h), col = NA, alpha = 0.2) +
  geom_line() +
  facet_grid(variable ~ type, scales = "free_y") +
  theme_bw() +
  scale_color_discrete(guide = "none") +
  scale_fill_discrete(guide = "none") +
  theme(axis.title = element_blank()) +
  geom_vline(xintercept = as_date(c("23/8/1", "23/12/1")))
```

```{r}
#| message: false
#| warning: false
read_tsv("outputs/tomst_daily_cleaned.tsv") %>% 
  filter(metric == "m") %>% 
  mutate(var_type = paste0(variable, " ", type)) %>% 
  select(tomst_sensornum, date, var_type, value) %>% 
  pivot_wider(names_from = var_type, values_from = value) %>% 
  select(-date, -tomst_sensornum) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  corrplot::corrplot(order = 'AOE', method = "number", type = "upper", 
                     number.digits = 1, number.cex = 0.8)
```

> Based on high correlations, I will focus first on daily values of of swc and t_air only.

## Dry season

According to the Guyaflux data, we defined the dry season as fortnightly under 50mm per month, from the 2023-07-22 to the 2023-11-7. Moreover, 2024 is particularly dry so we will focus on the 2023 wet season before the 2023-07-22.

```{r dsdef}
#| message: false
#| warning: false
readxl::read_xlsx("data/TourAFlux-2023-2024.xlsx") %>% 
  mutate(date = as_date(paste0(Year, "-1-1"))+Day-1, pr = `Somme de Pluie`) %>% 
  select(date, pr) %>% 
  group_by(date = floor_date(date, "15 days")) %>% 
  summarise(pr = sum(pr, na.rm = TRUE)) %>% 
  ggplot(aes(date, pr)) +
  geom_rect(xmin = as_date('2023-7-22'), xmax = as_date('2023-11-7'), 
            fill  = "#fff4e0", ymin = 0, ymax = 50, alpha = 0.3) +
  geom_col(fill = "lightblue", width = 4) +
  theme_bw() +
  geom_hline(yintercept = 50, linetype = "dashed") +
  theme(axis.title = element_blank())
```
