```{r set}
#| message: false
#| include: false
library(tidyverse)
```

# Macroclimate {.unnumbered}

## ERA5-Land

@muñoz-sabater2021

```{r era}
#| message: false
#| warning: false
#| fig-cap: "ERA5-Land raw temperature data for Paracou."
era <- read_tsv("data/raw_data/macro/era5.tsv")
g_year <- era %>% 
  group_by(year = year(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(year, tas)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank())
g_month <- era %>% 
  group_by(month = month(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(month, tas)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A",
                                "S", "O", "N", "D"))
g_day <- era %>% 
  group_by(minute = minute(time)+60*hour(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(minute/60, tas)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank())
cowplot::plot_grid(g_year, g_month, g_day)
```

## MeteoFrance

We computed back the correction of @hes2025 (see section 2 in SI) to determine a and b such $T_{ec} = \frac{T_e-b}{a}$ in order to correct ERA5-Land with Meteo France weather station data from Kourou CSG.

```{r weather_rel}
#| message: false
#| warning: false
#| fig-cap: "The temperature relationship between the hourly temperature at the weather station and the hourly temperature at ERA5-Land, as well as the correction factors a and b, following @hes2025.."
weather <- list.files("data/raw_data/macro/", pattern = "gz", full.names = TRUE) %>%
  read_delim(locale = locale(decimal_mark = "."), delim = ";") %>% 
  select(NOM_USUEL, LAT, LON, AAAAMMJJHH, RR1, TN50) %>%
  rename_all(tolower) %>%
  rename(loc = nom_usuel, pr = rr1, tas = tn50) %>%
  na.omit() %>% 
  mutate(aaaammjjhh = as.character(aaaammjjhh)) %>%
  mutate(year = str_sub(aaaammjjhh, 1, 4), 
         month = str_sub(aaaammjjhh, 5, 6),
         day = str_sub(aaaammjjhh, 7, 8),
         hour = str_sub(aaaammjjhh, 9, 10)) %>%
  mutate(time = as_datetime(paste0(year, "-", month, "-", day, 
                                   " ", hour, ":00:00"))) %>%
  select(loc, lat, lon, time, tas) %>% 
  filter(loc == "KOUROU CSG") %>% 
  left_join(read_tsv("data/raw_data/macro/era5_station.tsv") %>% 
              select(-lon, -lat),
            suffix = c("_station", "_era"), by = "time") %>% 
  na.omit()
a <- sd(weather$tas_era)/sd(weather$tas_station)
b <- mean(weather$tas_era)-a*mean(weather$tas_station)
weather %>% 
  ggplot(aes(tas_station, tas_era)) +
  geom_abline() +
  geom_point(alpha = .1, aes(col = hour(time))) +
  theme_bw() +
  coord_equal() +
  geom_smooth(method = "lm", se = FALSE) +
  ggpubr::stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
  ) +
  scale_color_viridis_c("Hour of the day") +
  xlab("Weather station hourly temperacture [°C]") +
  ylab("ERA5-Land hourly temperacture [°C]") +
  ggtitle(paste0("a = ", round(a, 2), " | b = ", round(b, 2)))
```

> The result differ from @hes2025 where $a=0.65$ and $b=8.72$. ⚠️ this correction as a strong effect on the resulting microclimatic buffer.

```{r weather}
#| message: false
#| warning: false
#| fig-cap: "MeteoFrance and ERA5-Land raw temperature data for Kourou CSG."
g_year <- weather %>% 
  mutate(tas_era_cor = (tas_era-13.5)/0.48) %>% 
  gather(type, tas, -loc, -lon, -lat, -time) %>% 
  group_by(type, year = year(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(year, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank(),
        legend.position = "bottom")
g_month <- weather %>% 
  mutate(tas_era_cor = (tas_era-b)/a) %>% 
  gather(type, tas, -loc, -lon, -lat, -time) %>% 
  group_by(type, month = month(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(month, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank(),
        legend.position = "bottom") +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A",
                                "S", "O", "N", "D"))
g_day <- weather %>% 
  mutate(tas_era_cor = (tas_era-b)/a) %>% 
  gather(type, tas, -loc, -lon, -lat, -time) %>% 
  group_by(type, minute = minute(time)+60*hour(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(minute/60, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank(),
        legend.position = "bottom")
cowplot::plot_grid(g_year, g_month, g_day)
```

## Debiasing

```{r}
#| message: false
#| warning: false
#| fig-cap: "ERA5-Land raw and debiased temperature data for Paracou."
era <- read_tsv("data/raw_data/macro/era5.tsv") %>% 
  mutate(tas_cor = (tas-b)/a) %>% 
  rename(raw = tas, debiased = tas_cor) %>% 
  gather(type, tas, -time, -lon, -lat)
g_year <- era %>% 
  group_by(type, year = year(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(year, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank())
g_month <- era %>% 
  group_by(type, month = month(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(month, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A",
                                "S", "O", "N", "D"))
g_day <- era %>% 
  group_by(type, minute = minute(time)+60*hour(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(minute/60, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank())
cowplot::plot_grid(g_year, g_month, g_day)
```
