```{r set}
#| message: false
#| include: false
library(tidyverse)
library(microclimr)
```

# Macroclimate {.unnumbered}

## ERA5-Land

We retrieved macroclimatic data from ERA5-Land [@muĂ±oz-sabater2021] at both Paracou and at the nearest weather station from Kourou space center (CSG).

```{r era}
#| message: false
#| warning: false
#| fig-cap: "ERA5-Land raw temperature data for Paracou."
era <- read_tsv("data/raw_data/macro/era5.tsv")
g_month <- era %>% 
  group_by(month = month(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(month, tas)) +
  geom_rect(xmin = 8, xmax = 12,
            ymin = -Inf, ymax = Inf, alpha = 0.5, fill  = "#fff4e0") +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A",
                                "S", "O", "N", "D"))
g_day <- era %>% 
  group_by(minute = minute(time)+60*hour(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(minute/60, tas)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank())
cowplot::plot_grid(g_month, g_day)
```

## Debiaising @hes2025

We computed back the correction of @hes2025 (see section 2 in SI) to determine a and b such $T_{ec} = \frac{T_e-b}{a}$ in order to correct ERA5-Land with Meteo France weather station data from Kourou CSG at hourly time step.

```{r weather_rel}
#| message: false
#| warning: false
#| fig-cap: "The temperature relationship between the hourly temperature at the weather station and the hourly temperature at ERA5-Land, as well as the correction factors a and b, following @hes2025.."
weather <- list.files("data/raw_data/macro/", pattern = "gz", full.names = TRUE) %>%
  read_delim(locale = locale(decimal_mark = "."), delim = ";") %>% 
  select(NOM_USUEL, LAT, LON, AAAAMMJJHH, RR1, TN50) %>%
  rename_all(tolower) %>%
  rename(loc = nom_usuel, pr = rr1, tas = tn50) %>%
  na.omit() %>% 
  mutate(aaaammjjhh = as.character(aaaammjjhh)) %>%
  mutate(year = str_sub(aaaammjjhh, 1, 4), 
         month = str_sub(aaaammjjhh, 5, 6),
         day = str_sub(aaaammjjhh, 7, 8),
         hour = str_sub(aaaammjjhh, 9, 10)) %>%
  mutate(time = as_datetime(paste0(year, "-", month, "-", day, 
                                   " ", hour, ":00:00"))) %>%
  select(loc, lat, lon, time, tas) %>% 
  filter(loc == "KOUROU CSG") %>% 
  left_join(read_tsv("data/raw_data/macro/era5_station.tsv") %>% 
              select(-lon, -lat),
            suffix = c("_station", "_era"), by = "time") %>% 
  na.omit()
a <- sd(weather$tas_era)/sd(weather$tas_station)
b <- mean(weather$tas_era)-a*mean(weather$tas_station)
weather %>% 
  ggplot(aes(tas_station, tas_era)) +
  geom_abline() +
  geom_point(alpha = .1, aes(col = hour(time))) +
  theme_bw() +
  coord_equal() +
  geom_smooth(method = "lm", se = FALSE) +
  ggpubr::stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~")),
  ) +
  scale_color_viridis_c("Hour of the day") +
  xlab("Weather station hourly temperacture [Â°C]") +
  ylab("ERA5-Land hourly temperacture [Â°C]") +
  ggtitle(paste0("a = ", round(a, 2), " | b = ", round(b, 2)))
```

> The result differ from @hes2025 where $a=0.65$ and $b=8.72$. âš ď¸Ź this correction as a strong effect on the resulting microclimatic buffer.

However the result is good to remove hour bias but increases bias at monthly and yearly timesteps.

```{r weather}
#| message: false
#| warning: false
#| fig-cap: "MeteoFrance and ERA5-Land raw temperature data for Kourou CSG."
g_month <- weather %>% 
  mutate(tas_era_cor = (tas_era-b)/a) %>% 
  gather(type, tas, -loc, -lon, -lat, -time) %>% 
  group_by(type, month = month(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(month, tas, col = type)) +
  geom_rect(xmin = 8, xmax = 12, group = NA, col = NA,
            ymin = -Inf, ymax = Inf, alpha = 0.5, fill  = "#fff4e0") +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank(),
        legend.position = "bottom") +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A",
                                "S", "O", "N", "D"))
g_day <- weather %>% 
  mutate(tas_era_cor = (tas_era-b)/a) %>% 
  gather(type, tas, -loc, -lon, -lat, -time) %>% 
  group_by(type, minute = minute(time)+60*hour(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(minute/60, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank(),
        legend.position = "bottom")
cowplot::plot_grid(g_month, g_day)
```

## Debiaising Fourier - CSG

The problem with the debiasing of @hes2025 is that it depends on the time-frequency used. We therefore also attempted debiasing using the relationship between the Fourier coefficients after decomposition. We will use the full time series for the Fourier transform, but missing data cannot be allowed for. Our focus will be on the microclimate study years of 2023 and 2024. The weather station data only miss nine measurements on six different days; we will therefore simply fill these in using local linear interpolation. This gives us 365 and 366 days, multiplied by 24 hours (as 2024 is a leap year). This produced a very large power spectrum over two years, with most power at periods of 0, 24 hours, 12 hours and their harmonics, as well as between periods of two years and one day. Looking at the ratio of power between ERA and the weather station, we can see that, as the frequency increases, the ERA power is smaller than that observed. This is logical given that a reanalysis is expected to smooth out small temporal variations. Nonetheless, even at low frequencies and long periods above the day (24-hour period), the power of the weather station tends to be higher than the power of the ERA5-Land reanalysis. We can use this computed ratio for the Kourou CSG station to interpolate the debiased ERA5-Land back at the Paracou station for the years 2023 and 2024.

```{r missing}
#| message: false
#| warning: false
#| fig-cap: "Missing data from the MeteoFrance temperature data from Kourou CSG."
join <- read_tsv("data/raw_data/macro/era5_station.tsv") %>% 
  arrange(time) %>% 
  filter(time >= as_datetime("2023-04-01 00:00:00")) %>% 
  filter(time <= as_datetime("2025-09-30 23:00:00")) %>% 
  rename(tas_era = tas) %>% 
  select(-lon, -lat) %>% 
  left_join(select(weather, time, tas_station)) %>% 
  mutate(date = as_date(time))
missing <- filter(join, is.na(tas_station)) %>% 
  select(date) %>% 
  unique()
join %>% 
  arrange(time) %>%
  mutate(tas_approx = zoo::na.approx(tas_station)) %>% 
  filter(date %in% missing$date) %>% 
  ggplot(aes(x = hour(time))) +
  geom_point(aes(y = tas_era, col = "era")) +
  geom_point(aes(y = tas_approx, col = "approx")) +
  geom_point(aes(y = tas_station, col = "station")) +
  facet_wrap(~ date) +
  theme_bw() +
  xlab("") +
  ylab("Temperature [Â°C]") 
```

```{r}
#| message: false
#| warning: false
#| fig-cap: "Caption."
join %>% 
  filter(time < as_datetime("2023-04-07 23:00:00")) %>% 
  ungroup() %>% 
  ggplot(aes(time)) +
  geom_line(aes(y = tas_era, col = "ERA5-Land")) +
  geom_line(aes(y = tas_station, col = "MeteoFrance weather station")) +
  theme_bw() +
  xlab("") + 
  ylab("Temperature [Â°C]") +
  ggtitle("Centre Spatial Guyanais (CSG)") +
  theme(legend.position = "bottom") +
  scale_color_discrete("")
```

```{r ps}
#| message: false
#| warning: false
#| fig-cap: "Power spectrums across 1-year for ERA5-Land and MeteoFrance temperature data from Kourou CSG."
duration <- length(unique(join$date))
fc <- join %>% 
  arrange(time) %>%
  mutate(tas_station = zoo::na.approx(tas_station)) %>% 
  select(-date) %>% 
  gather(source, tas, -time) %>% 
  mutate(source = gsub("tas_", "", source)) %>% 
  group_by(source) %>% 
  do(fft = fft_tab(.$tas, 24*duration)) %>%
  unnest(fft)
g_sub <- fc %>% 
  filter(period != 0) %>% 
  filter(period >= 12) %>% 
  ggplot(aes(frequency, power, fill = source)) +
  geom_col(position = position_dodge()) +
  theme_bw() +
  xlab("Period [h]") +
  ylab("Power") +
  scale_x_continuous(
    breaks = c(0, 1/(30*24), 1/(7*24), 1/48, 1/24, 1/12),
    labels = c("0", "1m", "7d", "48", "24", "12")
  ) +
  theme(legend.position = "none", 
        axis.title = element_blank())
fc %>% 
  filter(period != 0) %>% 
  ggplot(aes(frequency, power, fill = source)) +
  geom_col(position = position_dodge()) +
  theme_bw() +
  xlab("Period [h]") +
  ylab("Power") +
  scale_x_continuous(
    breaks = c(0, 1 / 24, 1 / 12, 1 / 8, 1 / 6),
    labels = c("0", "24", "12", "8", "6")
  ) +
  theme(legend.position = "bottom") +
  annotation_custom(ggplotGrob(g_sub),
                    ymin = 1, ymax = 3.5,
                    xmin = 1/12, xmax = 1/2) +
  annotate("text", x=1/24, y=2, label = paste0("Mean:", round(fc$power[1],2)))
```

```{r ratios}
#| message: false
#| warning: false
#| fig-cap: "Power spectrums ratios across 1-year for ERA5-Land and MeteoFrance temperature data from Kourou CSG."
ratios <- fc %>% 
  mutate(amplitude = Mod(coefficient)) %>% 
  select(-power, -coefficient) %>% 
  pivot_wider(names_from = source, values_from = amplitude) %>% 
  mutate(ratio = station/era)
ratios %>% 
  ggplot(aes(frequency, log(ratio))) +
  geom_point() +
  theme_bw() +
  scale_x_continuous(
    breaks = c(0, 1/(7*24), 1/48, 1/24, 1/12, 1/8, 1/6),
    labels = c("0", "Week", "48", "24", "12", "8", "6")
  ) +
  xlab("Period [h]") +
  ylab("Mod of the ratio of coefficients for Station/Era") +
  geom_smooth(method = "lm", se = FALSE) +
    ggpubr::stat_regline_equation(
    aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))
  )
```

## Debiaising Fourier - Paracou

Using adjustment values between Fourier coefficients of the weather station and ERA5-Land we can then de-bias ERA5-Land at the Paracou station. This resulted in higher predicted variation of temperature than the one smoothed by ERA5-Land reanalysis. However note that the extremes might be a bit exaggerated.

```{r paracou_fourier_debias_1}
#| message: false
#| warning: false
era <- read_tsv("data/raw_data/macro/era5.tsv") %>% 
  filter(time >= as_datetime("2023-04-01 00:00:00")) %>% 
  filter(time <= as_datetime("2025-09-30 23:00:00")) %>% 
  select(-lon, -lat)
fc_cor <- fft_tab(era$tas, 24*duration) %>% 
  left_join(select(ratios, frequency, period, ratio)) %>% 
  mutate(coefficient_c = coefficient*ratio)
era$tas_corf <- fft_reconstruct(
  c = fc_cor$coefficient_c,
  f = fc_cor$frequency[-1], 
  time = (1:(24*duration))-1
)
era %>% 
  filter(time < as_datetime("2023-04-07 23:00:00")) %>% 
  gather(source, tas, -time) %>% 
  mutate(source = recode(source,
                         "tas" = "station",
                         "tas_corf" = "Fourier"
                         )) %>% 
  ggplot(aes(x = time, y = tas, col = source)) +
  geom_line() +
  theme_bw() +
  xlab("") + 
  ylab("Temperature [Â°C]") 
```

Interestingly, we find almost back the debaising of @hes2025 with huorly variation but now the monthly variations are not exaggerated as in @hes2025, which confirm the interest of the approach for a debiasing at multiple time frequencies.

```{r paracou_fourier_debias_2}
#| message: false
#| warning: false
#| fig-cap: "ERA5-Land raw and debiased temperature data for Paracou."
era <- read_tsv("data/raw_data/macro/era5.tsv") %>% 
  filter(time >= as_datetime("2023-04-01 00:00:00")) %>% 
  filter(time <= as_datetime("2025-09-30 23:00:00")) %>% 
  mutate(tas_cor = (tas-b)/a) %>% 
  rename(raw = tas, hes2025 = tas_cor) 
era$Fourier <- fft_reconstruct(
  c = fc_cor$coefficient_c,
  f = fc_cor$frequency[-1], 
  time = (1:(24*duration))-1
)
era <- era %>% 
  gather(type, tas, -time, -lon, -lat)
write_tsv(era, "data/derived_data/era5_cor.tsv")
g_month <- era %>% 
  group_by(type, month = month(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(month, tas, col = type)) +
  geom_rect(xmin = 8, xmax = 12, group = NA, col = NA,
            ymin = -Inf, ymax = Inf, alpha = 0.5, fill  = "#fff4e0") +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank()) +
  scale_x_continuous(breaks = 1:12,
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A",
                                "S", "O", "N", "D"))
g_day <- era %>% 
  group_by(type, minute = minute(time)+60*hour(time)) %>% 
  summarise(tas = mean(tas)) %>% 
  ggplot(aes(minute/60, tas, col = type)) +
  geom_line() +
  theme_bw() +
  theme(axis.title = element_blank())
cowplot::plot_grid(g_month, g_day, nrow = 2)
```
